Bayesian Hierarchical Models for Common Components across Multiple System Configurations

### 2017-12-13 paperDraft3.tex
update

% need to recheck these numbers %
\begin{tabular}{|l|l|r|}
\multicolumn{3}{c}{\textbf{Goodness of Fit}} \\
\cline{1-3}
Distribution    & Structure & DIC \\
\hline
Exponential   & All Common Rate                      & 20,778.09             \\
              & All Hierarchical Rate                & 21,308.72             \\
              & Mixture Rate                         & 20,606.75             \\
Weibull       & All Common Rate                      & 20,202.39             \\
              & All Hierarchical Rate                & 17,984.85             \\
              & Mixture Rate                         & \textbf{17,710.85}    \\
\hline
\end{tabular}

The mixture rate models use a combination of common rate and hierarchical rate
parameterizations depending on the failure mode.  While we have shown that the
hierarchical Weibull model fits much better than the exponential, these test
still do not tell us that this model is a good fit for the data.  The last
method we will present is called posterior predictive checking ~\cite{ref3}.
Here we are given important features of the data. In the JLTV case we were
interested in the total failure counts for each phase and the number of time the
miles between failures was less than 140.  We then used the final model to
simulate 5,000 new datasets.  We then plot histograms for each of the 8 vehicles
for all 3 phases.  In Figure 1 are examples of the histograms produced, with the
line showing where the value of the true dataset fell.  For this method we don’t
expect all of the true values to fall in the center of the distribution.  Values
in the tales are to be expected in a random processes like this one.  We will
only be concerned if we find true values falling in the extreme tails or if we
see a common bias to the high or low side of the distributions.

@@

\subsection{Assurance Test Planning}
In this section, we use the resulting reliability model from the previous
section develop a plan for future testing.  Here the objective is to demonstrate
that at a desired level of confidence, the system will meet or exceed a
specified requirement. The methods to be employed will be similar to the
assurance testing as discussed by Hamada et al. ~\cite{ref4}. Bayesian
\emph{reliability assurance tests} are used to ensure that the reliability of
an item meets or exceeds a specified requirement with a desired probability.
Meeker and Escobar ~\cite{ME04} established the difference between this and the
more traditional \emph{reliability demonstration test}.  The difference being
that the assurance test plan incorporates supplementary data and information,
where the demonstration test only uses test data.

In both test planning methods we are interested in controlling error rates while
minimizing the resources required for testing.  In the DoD acquisition process,
the error rates will be referenced in terms of risk. \emph{Consumer risk} which
considers the event of purchasing a product that does not meet reliability
requirements.  \emph{Producer risk} which considers the event of a product with
acceptable reliability failing a given test and not being considered for
purchase.  Introducing the notation for this section, In the JLTV case study, we
will be determining how many miles on test $T$, are required and how many system
failures to allow $c$, before the product is considered unacceptable.  We will
define $W(t)$ as a random variable that represents the total number of system
failures after $t$ miles.  Suppose that $\pi$ denotes some quantity of interest
that is related to system reliability at a given time. It is common to base both
classical and Bayesian test plans on two specified levels of $\pi$: $\pi_0$, an
\emph{acceptable reliability level} (ARL), and $\pi_1$, a \emph{rejectable
reliability level} (RRL), where $\pi_1 \leq  \pi_0$. Although the precise
definition of ARL and RRL differ between the classical and Bayesian test
criteria, we use them in an equivalent way.

\subsubsection{Classical Approach}
It is quite common to use two criteria in determining classical test plans. The
\emph{producer's risk}, defined as the probability of failing the test when $\pi =
\pi_0$, and the \emph{consumer's risk}, defined as the probability of passing
the test when $\pi = \pi_1$.  Suppose that we specify a maximum value, $\alpha$,
of the producer's risk and a maximum value, $\beta$, of the consumer's risk. For
test planning, these criteria become
$$
\begin{aligned}
	\text{Producer's Risk} &= P \text{(Test Is Failed } \vert \pi_0 \text{)} \\ &=
	P \text{(} W(t) > c \vert \pi_0 \text{)} \leq \alpha
\end{aligned}
$$
and
$$
\begin{aligned}
	\text{Consumer's Risk} &= P \text{(Test Is Passed } \vert \pi_1 \text{)} \\ &=
	P \text{(} W(t) \leq c \vert \pi_1 \text{)} \leq \beta
\end{aligned}
$$
To choose a test plan for specified values of $(\alpha, \pi_0, \beta, \pi_1)$,
we assume a distributional form that defines the relationship between the number
of system failyres $W(t)$ and the reliability $\pi$.  Then we can simply find
the combinations of these probabilities by simultaneously solving these two
equations.  In the case where the component failure times are assumed to be
exponentially distributed with rate parameter $\lambda_i$ the failure counts for
each component come from a homogeneous Poisson process $w_i(t) \sim
\text{Poisson}(\lambda_i t)$, and the system failure counts $W(t) \sim
\text{Poisson}(\lambda_S t)$.  Numerous textbooks provide additional details of
this purely classical approach.

@@

\subsubsection{Weibull Case}
In the exponential case, we defined our quantity of interest that is related to
system reliability at a given time $\pi$ to be equal to the system failure rate
$\lambda_S$.  This distributional assumption gave us a convenient distributional
form for our system failure counts with $W(t) \sim \text{Poisson}(\lambda_S t)$
When we assume the miles to failure for each system component follows a Weibull
distribution the system failure counts no longer follow a homogeneous Poisson
process.

% something about how using bad as old makes this work...
Because each component failure rate now varies with time each
component failure count follows a non-homogeneous Poisson process $w_i(t) \sim
\text{Poisson}(\gamma_i \lambda_i t^{\gamma_i})$.  In this case we no longer
have a closed form for the system failure rate $\lambda_S$.

% need to redo this without the single rate %
The exception being
if we can assume a single shape parameter $\gamma$ for all components.  In this
case, it can be shown [A1] that the system failure count follows a non-homogeneous
Poisson process $W(t) \sim \text{Poisson}(\gamma \lambda_S t^{\gamma})$.  Now,
we can use the same approach as with the exponential distribution. An expression
for the posterior producer's risk for the exponential test plan $(T,c)$ is
$$
\begin{aligned}
    P(\lambda_S \geq \lambda_0 \; \vert \; \text{Test Is Failed)} &= P(\lambda_S
    \geq \lambda_0 \; \vert \; W > c) \\ &= \int_{\lambda_0}^{\infty}
    p(\lambda_S \; \vert W > c) d\lambda_S \\ &= \int_{0}^{\infty}
    \int_{\lambda_0}^{\infty} \frac{f(W > c \vert \lambda_S)}{\int_{0}^{\infty}
    \int_{0}^{\infty} f(W > c \vert \lambda_S) d\lambda_S d\gamma} d\lambda_S
    d\gamma \\ &= \frac{\int_{0}^{\infty} \int_{\lambda_0}^{\infty} 1 -
    [\sum_{W=0}^c \frac{(\gamma \lambda_S T^{\gamma})^W exp(-\gamma \lambda_S
    T^{\gamma})}{W!}]p(\lambda_S)p(\gamma)d\lambda_S d\gamma} {\int_{0}^{\infty}
    \int_{0}^{\infty} 1 - [ \sum_{W=0}^c \frac{(\gamma \lambda_S T^{\gamma})^W
    exp(-\gamma \lambda_S T^{\gamma})} {W!}]p(\lambda_S) p(\gamma)
    d\lambda_Sd\gamma} \leq \alpha
\end{aligned}
$$
\\
\\
Once again we can simplify by setting c to be zero and then perform Monte Carlo
integration using N posterior draws for both $\lambda_S^{(j)}$ and
$\gamma^{(j)}$
$$
\begin{aligned}
	 P(\lambda_S \geq \lambda_0 \; \vert \; W = 0) &= \frac{\int_{0}^{\infty}
	 \int_{\lambda_0}^{\infty} 1 - exp(- \gamma \lambda_S T^{\gamma})p(\lambda_S)
	 p(\gamma) d\lambda_S d\gamma} {\int_{0}^{\infty} \int_{0}^{\infty} 1 - exp(-
	 \gamma \lambda_S T^{\gamma})p(\lambda_S) p(\gamma) d\lambda_S d\gamma} \\
	 &\approx \frac{\sum_{j = 1}^{N} 1 - exp(-\gamma^{(j)} \lambda_S^{(j)}
	 T^{\gamma^{(j)}}) I(\lambda_S^{(j)} \geq \lambda_0)} {\sum_{j = 1}^{N} 1 -
	 exp(- \gamma^{(j)} \lambda_S^{(j)} T^{\gamma^{(j)}})} \end{aligned}
$$

In the case of multiple shape parameters, these calculations can still be done.
The practitioner must be careful in defining the system reliability quantity of
interest, accounting for each component failure rate.  The computation can
quickly become  unmanagable as the number of components and/or number of allowed
failures becomes large.

@@

\section{Appendix A1}
\subsection{Series System Failure Rate}

\textbf{Definitions:} \\
\noindent
Reliability: $R(t) = 1 - F(t)$ \\
Hazard or failure rate function: $\lambda(t) = \frac{f(t)}{R(t)}$ \\
Series system reliability: $R_S(t) = \prod_{i = 1}^N R_i(t)$
\\
\\
\textbf{Result:}
$$
\begin{aligned}
	R_S(t) &= \prod_{i = 1}^N R_i(t) \\
	\frac{dR_S(t)}{dt} &= \frac{d}{dt} \prod_{i = 1}^N R_i(t) \; \; \text{(derivative of both sides)} \\
	\frac{dR_S(t)}{dt} &= \sum_{i=1}^N \left[ \frac{dR_i(t)}{dt} \frac{dR_S(t)}{dR_i(t)} \right] \; \; \text{(product rule)} \\
    \frac{-\frac{dR_S(t)}{dt}}{R_S(t)} &= \sum_{i=1}^N \left[  \frac{-\frac{d}{dt}R_i(t)}{dR_i(t)} \right] \; \; \text{(divide both sides by $-R_S(t)$)} \\
    \lambda_S(t) &= \sum_{i = 1}^N \lambda_i(t) \; \; \text{(because $\lambda(t) = \frac{f(t)}{R(t)} = \frac{-\frac{dR(t)}{dt}}{R(t)}$ )}
\end{aligned}
$$

\subsection{Non-Homogeneous Poisson Process}

\textbf{Definitions:} \\
\noindent
Number of failures before time t: $N(t) \sim \text{Poisson}(m(t))$ \\
Mean function: $m(t) = \int_0^t \lambda(s)ds$ (represents the expected number of failures before time t) \\
Weibull($\gamma, \beta $) failure rate: $\lambda(t) = \gamma\beta t^{\beta - 1}$
\\
\\
\textbf{Result:}\\
If we assume a constant shape parameter $\beta$ then,\\
$$
\begin{aligned}
	\lambda_S(t) &= \sum_{i = 1}^N \lambda_i(t) \\
    &= \sum_{i = 1}^N \gamma_i\beta t^{\beta - 1} \\
    &= \beta t^{\beta -1} \sum_{i = 1}^N \gamma_i
\end{aligned}
$$
\newpage
Then Solving for the mean function of the system,

$$
\begin{aligned}
	m_S(t) &= \int_0^t \lambda_S(s)ds \\
    &= \int_0^t \beta t^{\beta -1} \sum_{i = 1}^N \gamma_i \\
    &= \sum_{i = 1}^N \gamma_i \int_0^t \beta t^{\beta -1} \\
    &= t^\beta \sum_{i = 1}^N \gamma_i
\end{aligned}
$$

This gives us: $N(t) \sim \text{Poisson}(t^\beta \sum_{i = 1}^N \gamma_i)$

\subsection{Notes}
Need to say somewhere we are assuming "bad-a-old" for the Weibull model.

>>>>

% need to recheck these numbers %
\begin{tabular}{|l|l|r|}
\multicolumn{3}{c}{\textbf{Goodness of Fit}} \\
\cline{1-3}
Distribution    & Structure & DIC \\
\hline
Exponential   & Common Rate Model                     & 21,264.19             \\
              & Hierarchical Rate Model               & 21,244.47             \\
Weibull       & Common Rate Model                     & 20,658.25             \\
              & Hierarchical Rate Model               & \textbf{17,908.06}    \\
\hline
\end{tabular}

Here we have shown that the hierarchical Weibull model fits the data much better
than the exponential, but in general information criteria comparisons do not
tell us if the model is truely a good representation of the underlying process.
The last method we will present is called posterior predictive checking
~\cite{ref3}. Here we are given important features of the data. In the JLTV case
we were interested in the total failure counts for each phase and the number of
time the miles between failures was less than 140.  We then used the final model
to simulate 5,000 new datasets. We then plot histograms for each of the 8
vehicles for all 3 phases.  In Figure 1 are examples of the histograms produced,
with the line showing where the value of the true dataset fell.  For this method
we don’t expect all of the true values to fall in the center of the
distribution.  Values in the tales are to be expected in a random processes like
this one.  We will only be concerned if we find true values falling in the
extreme tails or if we see a common bias to the high or low side of the
distributions.

@@

\subsection{Assurance Test Planning}
In this section, we will use the reliability models fit in section two section
to develop a plan for future testing.  Here the objective is to demonstrate that
at a desired level of confidence, the system will meet or exceed a specified
requirement. The methods to be employed will be similar to the assurance testing
as discussed by Hamada et al. ~\cite{ref4}. Bayesian \emph{reliability assurance
tests} are used to ensure that the reliability of an item meets or exceeds a
specified requirement with a desired probability. Meeker and Escobar
~\cite{ME04} established the difference between this and the more traditional
\emph{reliability demonstration test}.  The difference being that the assurance
test plan incorporates supplementary data and information, where the
demonstration test only uses test data.

In both test planning methods we are interested in controlling error rates while
minimizing the resources required for testing.  In the DoD acquisition process,
the error rates will be referenced in terms of risk. \emph{Consumer risk} which
considers the event of purchasing a product that does not meet reliability
requirements.  \emph{Producer risk} which considers the event of a product with
acceptable reliability failing a given test and not being considered for
purchase.  Introducing the notation for this section, in the JLTV case study, we
will be determining how many miles on test $T$, are required and how many system
failures to allow $c$, before the product is considered unacceptable.  We will
define $W(t)$ as a random variable that represents the total number of system
failures after $t$ miles.  We will let $\pi$ denote some quantity of interest
that is related to system reliability at a given time. It is common to base both
classical and Bayesian test plans on two specified levels of $\pi$: $\pi_0$, an
\emph{acceptable reliability level} (ARL), and $\pi_1$, a \emph{rejectable
reliability level} (RRL), where $\pi_1 \leq  \pi_0$. Although the precise
definition of ARL and RRL differ between the classical and Bayesian test
criteria, we use them in an equivalent way.

\subsubsection{Classical Approach}
In the frequentist (classical) version we will use two criteria in determining
the test plans. The \emph{producer's risk}, defined as the probability of
failing the test when $\pi = \pi_0$, and the \emph{consumer's risk}, defined as
the probability of passing the test when $\pi = \pi_1$.  Suppose that we specify
a maximum value, $\alpha$, of the producer's risk and a maximum value, $\beta$,
of the consumer's risk. For test planning, these criteria become
$$
\begin{aligned}
	\text{Producer's Risk} &= P \text{(Test Is Failed } \vert \pi_0 \text{)} \\ &=
	P \text{(} W(t) > c \vert \pi_0 \text{)} \leq \alpha
\end{aligned}
$$
and
$$
\begin{aligned}
	\text{Consumer's Risk} &= P \text{(Test Is Passed } \vert \pi_1 \text{)} \\ &=
	P \text{(} W(t) \leq c \vert \pi_1 \text{)} \leq \beta
\end{aligned}
$$
To choose a test plan for specified values of $(\alpha, \pi_0, \beta, \pi_1)$,
we assume a distributional form that defines the relationship between the number
of system failyres $W(t)$ and the reliability $\pi$.  Then we can simply find
the combinations of these probabilities by simultaneously solving these two
equations.  In the case where the component failure times are assumed to be
exponentially distributed with rate parameter $\lambda_i$ the failure counts for
each component come from a homogeneous Poisson process $w_i(t) \sim
\text{Poisson}(\lambda_i t)$, and the system failure counts $W(t) \sim
\text{Poisson}(\lambda_S t)$.  Numerous textbooks provide additional details of
this purely classical approach.

@@

\subsubsection{Weibull Case}
In the exponential case, we defined our quantity of interest that is related to
system reliability at a given time $\pi$ to be equal to the system failure rate
$\lambda_S$.  This distributional assumption gave us a convenient distributional
form for our system failure counts with $W(t) \sim \text{Poisson}(\lambda_S t)$
When we assume the miles to failure for each system component follows a Weibull
distribution the system failure counts no longer follow a homogeneous Poisson
process.  When assuming "bad as old" intra-test repairs we do find that the
failure count for a given failure mode does follow a nonhomogeneous Poisson
process, $w(t)_i \sim \text{Poisson}( (\lambda_i t)^{\gamma_i} )$.  In this case
we no longer have a closed form for the system failure rate $\lambda_S = \sum_i
(\lambda_i t)^{\gamma_i}$.  Because of this, using our previous approach leads to
an unpleasant series of integals with parameters in the bounds.  Thankfully we
can once again lean on Monte Carlo integration and the expression for the
posterior producer's risk for the Weibull test plan $(T,c)$ simplifies to
$$
\begin{aligned}
	 P(\lambda_S \geq \lambda_0 \; \vert \; W = 0) &= \frac{\int_{0}^{\infty}
	 \int_{\lambda_0}^{\infty} 1 - exp(- \gamma \lambda_S T^{\gamma})p(\lambda_S)
	 p(\gamma) d\lambda_S d\gamma} {\int_{0}^{\infty} \int_{0}^{\infty} 1 - exp(-
	 \gamma \lambda_S T^{\gamma})p(\lambda_S) p(\gamma) d\lambda_S d\gamma} \\
	 &\approx \frac{\sum_{j = 1}^{N} 1 - exp(-\gamma^{(j)} \lambda_S^{(j)}
	 T^{\gamma^{(j)}}) I(\lambda_S^{(j)} \geq \lambda_0)} {\sum_{j = 1}^{N} 1 -
	 exp(- \gamma^{(j)} \lambda_S^{(j)} T^{\gamma^{(j)}})}
\end{aligned}
$$
\\
Once again set c to be zero and used N posterior draws for both
$\lambda_S^{(j)}$ and $\gamma^{(j)}$

@@

\section{Appendix A1}
\textbf{Exponential Model:}
\begin{equation*}
L(\lambda_1, ..., \lambda_8, \rho^{P2}, \rho^{P3}, \vert Y) = \prod_{i = 1}^8
  \Bigg[\prod_{j = 1}^{n_j}(\lambda_i p^*e^{-\lambda_i p^* y_{ij}})\prod_{j =
  1}^R(e^{-\lambda_i p^*
  y_{ij}})\Big(\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda_i^{\alpha -
  1}e^{-\beta\lambda_i}\Big)\Bigg]
\end{equation*}
\[p^* =   \left\{
\begin{array}{ll}
      1 & y_{ij}\in \text{Phase 1} \\
      \rho_{j}^{P2} & y_{ij} \in \text{Phase 2} \\
      \rho_{j}^{P2} \rho_{j}^{P3} & y_{ij} \in \text{Phase }
\end{array}
\right. \]
$R$ is the number of right-censored observations, and n is the number of
non-censored observations.

\begin{equation*}
\pi(\alpha, \beta, \rho^{P2}, \rho^{P3}) \propto (\alpha^{a_1-1}e^{-\alpha
  a_2})(\beta^{b_1 - 1}e^{-\beta b_2})(\rho^{P2*c_1 - 1}e^{-\rho^{P2}
  c_2})(\rho^{P3*d_1 - 1}e^{-\rho^{P3} d_2})
\end{equation*}

\begin{align*}
\lambda_i|\cdot &\sim \text{Gamma}(n_i + \alpha, \sum_{j=1}^{n_i}y_{ij}p^* + \beta) \\
\beta|\cdot &\sim \text{Gamma}(8\alpha + b_1, \sum{\lambda_i} + b_2) \\
\rho^{P2}|\cdot &\sim \text{Gamma}(N_{1,2} + c_1, \sum{y_{ij}\lambda_i} + c_2) \\
\rho^{P3}|\cdot &\sim \text{Gamma}(N_{2} + d_1, \sum{y_{ij}\lambda_i} + d_2) \\
\alpha|\cdot &\sim \text{Unknown form.  Used Metropolis-Hastings.}
\end{align*}

$N_1$ and $N_2$ are the total non-censored observations for phases two and three
and phase three, respectively.

\textbf{Weibull Model (Bad as Old):}
\begin{equation*}
  L(\lambda_1, ..., \lambda_8, \gamma, \rho^{P2}, \rho^{P3}, \vert T) = \prod_{i =
    1}^8 \Bigg[\prod_{j = 1}^{n_j}(\lambda_i p^*t_{ij}^{\gamma_i-1} e^{-\lambda_i
    p^* t_{ij}^{\gamma_i}})\prod_{j = 1}^R(e^{-\lambda_i p^*
    t_{ij}^{\gamma_i}})\Big(\frac{\beta^\alpha}{\Gamma(\alpha)}\lambda_i^{\alpha -
    1}e^{-\beta\lambda_i}\Big)\Bigg]
\end{equation*}
\[p^* =   \left\{
\begin{array}{ll}
      1 & t_{ij}\in \text{Phase 1} \\
      \rho^{P2^\gamma} & t_{ij} \in \text{Phase 2} \\
      \rho^{P2^\gamma} \rho^{P3^\gamma} & t_{ij} \in \text{Phase }
\end{array}
\right. \]
$R$ is the number of right-censored observations, and n is the number of
non-censored observations.

\begin{equation*}
\pi(\alpha, \beta, \rho^{P2}, \rho^{P3}) \propto (\alpha^{a_1-1}e^{-\alpha
  a_2})(\beta^{b_1 - 1}e^{-\beta b_2})(\rho^{P2*c_1 - 1}e^{-\rho^{P2}
  c_2})(\rho^{P3 * d_1 - 1}e^{-\rho^{P3} d_2})(\gamma^{g_1 - 1}e^{-\gamma g_2})
\end{equation*}

\begin{align*}
\lambda_i|\cdot &\sim \text{Gamma}(n_i + \alpha, \sum_{j=1}^{n_i}t_{ij}^{\gamma_i} p^* + \beta) \\
\beta|\cdot &\sim \text{Gamma}(8\alpha + b_1, \sum{\lambda_i} + b_2) \\
\rho^{P2}|\cdot &\sim \text{Gamma}(N_{1,2} + c_1, \sum{t_{ij}^{\gamma_i} \lambda_i} + c_2) \\
\rho^{P3}|\cdot &\sim \text{Gamma}(N_{2} + d_1, \sum{t_{ij}^{\gamma_i} \lambda_i} + d_2) \\
\alpha|\cdot &\sim \text{Unknown form.  Used Metropolis-Hastings.}\\
\gamma|\cdot &\sim \text{Unknown form.  Used Metropolis-Hastings.}
\end{align*}

$N_1$ and $N_2$ are the total non-censored observations for phases two and three
and phase three, respectively.
